---
title: "Lab 4"
author: "Alex Thompson"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the data

```{r}
library(readxl)
defense <- read_excel('C:/Users/thomp/Documents/DS202/ds202_labs/ds202_labs/cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('C:/Users/thomp/Documents/DS202/ds202_labs/ds202_labs/cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)

biography <- read_excel('C:/Users/thomp/Documents/DS202/ds202_labs/ds202_labs/cyclonesFootball2018.xlsx', sheet='Biography')
str(biography)
```

## Cleaning the Data

```{r message=FALSE, warning=FALSE}
library(dplyr)

#Loading the dataframes in
defClean = defense
offClean = offense
bioClean = biography

#Converting name and opponent columns into factors for the three frames
defClean$Name = factor(defClean$Name)
defClean$Opponent_Opponent = factor(defClean$Opponent_Opponent)

offClean$Name = factor(offClean$Name)
offClean$Opponent_Opponent = factor(offClean$Opponent_Opponent)

bioClean$Name = factor(bioClean$Name)

#Changing columns into numeric in all three frames
defClean = mutate_at(defClean, vars(Tackles_Solo,Tackles_ASST,Tackles_TFL,Tackles_Sack,Turnover_FF,Turnover_FR,Turnover_INT,Pass_QBH,Pass_PB), funs(as.numeric))

offClean = mutate_at(offClean, vars(Receiving_REC,Receiving_YDS,Receiving_TD,Rushing_ATT,Rushing_YDS,Rushing_TD,'Passing_CMP-ATT',Passing_YDS,Passing_TD,Passing_INT), funs(as.numeric))

offClean = mutate_at(offClean, vars(Receiving_REC,Receiving_YDS,Receiving_TD,Rushing_ATT,Rushing_YDS,Rushing_TD,'Passing_CMP-ATT',Passing_YDS,Passing_TD,Passing_INT), funs(as.numeric))

bioClean = mutate_at(bioClean, vars(Weight), funs(as.numeric))

#Loading in the stringr package for string manipulation
library(stringr)

#Pulling feet and inches from height, putting them in their own respective columns
bioClean = bioClean %>%
   dplyr::mutate(inch = as.numeric(str_extract(Height, "[^-]+$")))

bioClean = bioClean %>%
   dplyr::mutate(feet = as.numeric(str_extract(Height, "[^-]+")))

#From feet and inch columns, recreate the Height column in terms of inches
bioClean = mutate(bioClean, Height = feet * 12 + inch)

#Dropping the feet and inch columns from the frame
bioClean = subset(bioClean, select = -c(feet, inch))

#Displaying the structure of the three data frames
str(defClean)

str(offClean)

str(bioClean)
```

## Tidying (A)

```{r}
library(tidyr)

#Reformating defClean such that Statistic is a key and the rest of the numeric are values
defClean = gather(defClean, key = Statistic, value = Value, 3:11)
```

```{r}
library(ggplot2)

#Drawing up a histogram of distribution of values, seperated by the Statistic type 

ggplot(defClean, aes(Value)) + geom_histogram(binwidth = .7) + facet_wrap(~Statistic) + labs(y="Count", title="Count of Values seperated by Statistic type")
```

**Interpretation:** Based on this visualization it seems like in general it is unlikely to see values that are remarkebly high, the distribtuions for all statistics are centered around lower values with very few outliers into the higher range of values. 

```{r message=FALSE, warning=FALSE}
#Spreading defClean like the red sea, just kidding, but I am spreading defensive statistics by which opponent that ISU was playing against
defClean = spread(defClean, key = Opponent_Opponent, value = Value)

#Now I am filtering to only have entries for solo tackles statistics
tackSol = filter(defClean, Statistic == "Tackles_Solo")

#Rendering a point graph based off of the cleaned data
ggplot(tackSol, aes(x=tackSol$`Kansas State`, y=tackSol$`West Virginia`)) + geom_point() + labs(x="Kansas State", y="West Virginia", title= "Solo Tackles against West Virginia vs. against Kansas State")
```

**Interpretation:** Based on this visualization it seems like palyers who were largely locked out by West Virginia were still able to score some solo tackles, or even many solo tackles against Kansas State. This leads me to the conclusion that ISU had better defense in net against Kansas State.

## Tidying (B)

```{r}
#Seperating the state column into town and state
bioClean = separate(bioClean, Hometown, c("town", "state"), sep = ", ")

#Show entry frequencies for each state
table(bioClean$state)
```

## Joining frames

```{r message=FALSE, warning=FALSE}
#Joining Offense and bioClean to show after each stat entry where the player is from
offense = left_join(offClean, bioClean, by = "Name")

#Calcualte correlation between weight and Recieving Yards
cor(offense$Weight, offense$Receiving_YDS, use = "complete.obs")
```

**Interpretation:** With such a low correlation value it is unlikely that there is any relationship between a players weight and their Recieving yards. 
